### 2020-04-13 reports/thesis/chapters/5_inference.tex
Final revise
Corrections
CpEmAlgorithm does not plot estimated std
roc calculates area exactly
Recoded NullIid and included biweighted mean

A pixel by pixel hypothesis test was proposed to do defect detection. Beforehand, linear shading correction was applied to the projections. The \emph{aRTist} projections were shading corrected using the simulated greyscale projections. The test statistic, for the pixel located at $(x,y)$, is
\begin{equation}
  Z_{x,y} = 
  \dfrac{
    \text{projection}_{x,y} - \emph{aRTist}_{x,y}
  }
  {
    \sqrt{\widehat{\variance}\left[\text{projection}_{x,y}\right]}
  }
\end{equation}
where $\widehat{\variance}\left[\text{projection}_{x,y}\right]$ is the estimated grey value variance. The test statistic was calculated for each pixel in the ROI, that is, pixels which represent the test sample. The ROI was created by manually segmenting the test sample from the projection.

For high photon rate, it was shown in the previous chapters that the grey values in the scan can be modelled using a Normal distribution. Thus, if there are no significant differences between the obtained and \emph{aRTist} projections, then the randomness of the test statistics can be approximately quantified as
\begin{equation}
Z_{x,y}\sim \normal(0,1) \ .
\end{equation}
As with usual statistical convention, an upper case $Z$ denotes a random variable. A lower case $z$ denotes a realisation or an observation of that random variable.

To illustrate this method of inference, the projections from the dataset \texttt{AbsFilter} at \ang{120} was used. To recap, replicate projections of a test sample, with purposefully manufactured voids, were obtained. \emph{aRTist} was used to create a simulation of that projection but as if the voids were not there. Thus, the method should pick these voids up.

The obtained and \emph{aRTist} projections are shown in Figures \ref{fig:inference_inferenceIntro_inference_scan} and \ref{fig:inference_inferenceIntro_inference_artist} respectively. The 20 replicate projections were split into two. 19 randomly selected projections were used for the variance-mean model to fit onto. A gamma GLM, with a basic linear relationship, was used for the variance-mean model. The remaining projection was compared with the \emph{aRTist} projection. The variance was predicted using the grey value in the \emph{aRTist} simulation as the predictor variable.

@@

@@

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$. The method used in fitting the smoothing spline could contribute to the underfitting. For example, if the smoothing spline was not smooth enough so that multiple local minima appeared, it would add variance to the optimal bandwidth if the optimisation algorithm converges to a local minimum. A basic linear relationship was chosen to avoid overfitting and to keep the rule of thumb simple.

@@

Various estimates of central tendency and dispersion were compared with the empirical null mean and empirical null standard deviation. The empirical sampling distribution of these estimators was observed to look for properties such as bias, variance and robustness. The central tendency estimates looked at were the median and the mean. The dispersion estimates looked at were the median absolute deviation (MAD) around the mode $\times 1.483$, interquartile range (IQR) $\div 1.349$ and the standard deviation. The normalisation constants were used so that the MAD and IQR can be compared directly to the standard deviation.

A dataset of $n = \pi r^2$ standard Normal random variables was simulated to obtain estimates of central tendency and dispersion. This was repeated 100 times to obtain a sampling distribution. The results are shown in Figures \ref{fig:inference_NullIidMean} and \ref{fig:inference_NullIidStd}. The empirical null mean and empirical null standard deviation both suffered from high variance, much higher than typical robust estimators such as the median and IQR.

@@

To test for robustness, a dataset of $n = \pi r^2$ mixture of Gaussian random variables were simulated with the following distribution: $\normal(0,1)$ with probability $90\%$, $\normal(3,1)$ with probability $10\%$. The $\normal(3,1)$ component acts as non-null statistics and robust estimators should be unaffected by it. The results are shown in Figures \ref{fig:inference_NullIidMixtureMean} and \ref{fig:inference_NullIidMixtureStd}. The empirical null estimators are robust to the non-null component but still suffers from high variance. The rest of the estimators were affected by the non-null statistics, making them unsuitable to estimate parameters of the null parameters.

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidEmpirical_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMedianIqr_nullMean.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMeanVar_nullMean.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMean}
\end{figure}

\begin{figure}
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidEmpirical_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMadMode_nullStd.eps}
        \caption{MAD mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMedianIqr_nullStd.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMeanVar_nullStd.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidStd}
\end{figure}

\begin{figure}
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureEmpirical_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMedianIqr_nullMean.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMeanVar_nullMean.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMixtureMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureEmpirical_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMadMode_nullStd.eps}
        \caption{MAD mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMedianIqr_nullStd.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMeanVar_nullStd.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidMixtureStd}
\end{figure}

@@

The empirical null filter extends the empirical null to images. The parameters of the null distribution varies spatially, slowly and smoothly. Let the test statistic of the pixel at position $(x,y)$ be
\begin{equation}
Z_{x,y}\sim\normal(\mu_{x,y},\widehat{\sigma}_{x,y}^2)
\end{equation}
for $x=1,2,\dotsc,W$ and $y=1,2,\dotsc,H$. $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ are the null mean and null standard deviation at position $(x,y)$ respectively. Define the null hypotheses to be
\begin{equation}
  H_{0,x,y}:\mu_{x,y}=\widehat{\mu}_{0,x,y}
\end{equation}
which are tested against
\begin{equation}
  H_{1,x,y}:\mu_{x,y}\neq\widehat{\mu}_{0,x,y} \ .
\end{equation}
 The empirical null filter aims to estimate the parameters $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ for all $x$ and $y$.

@@

The parameter of choice is the kernel radius $r$. A $r$ too small will treat defects as the null and will not detect them. A $r$ too big can also cause problems as well. The empirical null filter assumes that the null parameters vary smoothly and slowly so that the null test statistics captured by the kernel is somewhat Normal. If $r$ is too big, the kernel will capture distanced test statistics with very different null distributions. If all of these test statistics are pulled together into a histogram, the null distribution may not look Normal at all which can cause problems when fitting a Normal null distribution. Furthermore, if $r\rightarrow \infty$, then this is just using the empirical null on all pixels.

$r$ controls the proportion of pixels which are null captured by the kernel, denoted as $\pi_0$. $r$ should be selected so that $\pi_0$ is of sensible value, for example, $>90\%$. In practice, the defects are unknown so various kernel radiuses should be tried out. The kernel radius should be much larger than the size of the defect.

@@

An experiment was conducted to demonstrate the effects of the empirical null filter on a simulated $\SI{256}{\pixel}\times\SI{256}{\pixel}$ Gaussian image, where all the pixels have value distributed as standard Normal. The purpose is to simulate test statistics which are null and no positive results should be detected. The test statistics after filtering should preserve its original distribution, or at least close to it. The statistical moments before and after filtering should be the same and were investigated. Because the empirical null mean and empirical null standard deviation are random variables, the filtered pixels will never be Normal and any normality tests would be too strict for this experiment.

Contamination is defined as some smooth and slowly varying function added and/or multiplied to an image. Conducting hypotheses testing on a contaminated Gaussian image may cause errors if the null distributions were incorrectly specified as standard Normal. In this experiment, the contamination was a gradient such that the pixel at position $(x,y)$ have a value distributed as $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_{0,x,y}^2)$ where
\begin{equation}
  \mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0) \ ,
\end{equation}
$(x_0,y_0)$ is the centre of the image and $\sigma_{0,x,y}=2$ for all $(x,y)$.

A caveat is that some bias to the null standard deviation estimation would be introduced because of the contamination. This is because sources of variance captured by the kernel are from $\sigma_{0,x,y}$ and also from the variability of $\mu_{0,x,y}$. This can be shown. Let $Z_{x,y}$ be the value of the pixels captured by the circular kernel for integer values of $(x,y)$ such that $x^2+y^2\leqslant r^2$. Suppose that $Z_{x,y}\sim\normal(ax+by,c^2\sigma^2)$. The calculations of the expectation and variance of $Z_{x,y}$ can be approximated by treating $x$ and $y$ as uniformly distributed within a circle centred at the origin with radius $r$. The expectation is $\expectation[Z_{X,Y}] = \expectation\expectation[Z_{X,Y}|X,Y]=0$. The variance is given as $\variance[Z_{X,Y}] = \expectation\variance[Z_{X,Y}|X,Y] + \variance\expectation[Z_{X,Y}|X,Y]$. From the distribution of $Z_{X,Y}$, $\expectation\variance[Z_{X,Y}|X,Y] = c^2\sigma^2$ and $\variance\expectation[Z_{X,Y}|X,Y]=\variance[aX+bY]$. This is worked out to be
\begin{align}
\variance[aX+bY] &= \dfrac{1}{\pi r^2}\int_{\rho=0}^{\rho=r}\int_{\theta=0}^{\theta=2\pi}
\rho^3(a\cos\theta + b\sin\theta)^2 \diff \theta \diff \rho
\nonumber\\
& = \dfrac{1}{4}(a^2+b^2)r^2
\end{align}
thus
\begin{equation}
\variance[Z_{X,Y}] = \dfrac{1}{4}(a^2+b^2)r^2 + c^2\sigma^2 \ .
\end{equation}
This shows that when estimating the null variance on a contaminated Gaussian image, there may be a $(a^2+b^2)r^2/4$ bias. If the contamination is slowly varying, that is $a$ and $b$ are small, then the bias should be small.

@@

For a given kernel radius $r$, 100 different Gaussian images were filtered to investigate the within image mean, standard deviation and kurtosis of the filtered statistics. Various other filters for normalisation were used as well such as the MAD mode null filter, median IQR null filter and the mean variance null filter. They use different estimators to estimate the null parameters as their name suggests, Table \ref{table:inference_nullFilters} clarifies them.

\begin{table}
  \centering
  \begin{tabular}{l|l|l}
    Filters&Null mean&Null standard deviation\\\hline
    MAD mode null&Mode&MAD around the mode $\times1.483$\\
    Median IQR null&Median&IQR $\div1.349$\\
    Mean variance null&Mean&Standard deviation
  \end{tabular}
  \caption{Various filters using different estimators to estimate the null parameters are described here.}
  \label{table:inference_nullFilters}
\end{table}

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_mean.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_mean.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_mean.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_mean.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample mean of a filtered Gaussian image of size $\SI{256}{\pixel}\times\SI{256}{\pixel}$.  The boxplots summarise the 100 repeated simulations of the image. The dashed lines show the 95\% confidence interval using standard tests and assuming independence.}
  \label{fig:inference_AllNullGaussian_mean}
\end{figure}

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_variance.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_variance.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_variance.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_variance.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample standard deviation of a filtered Gaussian image of size $\SI{256}{\pixel}\times\SI{256}{\pixel}$. The boxplots summarise the 100 repeated simulations of the image. The dashed lines show the 95\% confidence interval using standard tests and assuming independence.}
  \label{fig:inference_AllNullGaussian_var}
\end{figure}

>>>>

A pixel by pixel hypothesis test was proposed to do defect detection. To illustrate this method of inference, the projections from the dataset \texttt{AbsFilter} at \ang{120} was used. To recap, 20 replicate projections of a test sample, with purposefully manufactured voids, were obtained. \emph{aRTist} was used to create a simulation of that projection but as if the voids were not there. Thus, the method should pick these voids up.

Beforehand, linear shading correction was applied to the projections. The \emph{aRTist} projections were shading corrected using the simulated greyscale projections. The 20 replicate projections were split into two. 19 randomly selected projections were used for the variance-mean model to fit onto. The remaining projection was compared with the \emph{aRTist} projection. This remaining projection and the \emph{aRTist} projections are shown in Figures \ref{fig:inference_inferenceIntro_inference_scan} and \ref{fig:inference_inferenceIntro_inference_artist} respectively.

A gamma GLM, with a basic linear relationship, was used for the variance-mean model, as described in Chapter \ref{chapter4}. The variance was predicted using the grey value in the \emph{aRTist} simulation as the predictor variable.

The test statistic, for the pixel located at $(x,y)$, is
\begin{equation}
  Z_{x,y} =
  \dfrac{
    \text{projection}_{x,y} - \emph{aRTist}_{x,y}
  }
  {
    \sqrt{s^2\left[\emph{aRTist}_{x,y}\right]}
  }
\end{equation}
where $s^2\left[\emph{aRTist}_{x,y}\right]$ is the predicted grey value variance. The test statistic was calculated for each pixel in the ROI, that is, pixels which represent the test sample. The ROI was created by manually segmenting the test sample from the projection.

It is important to identity which quantites are random and which are not. For high photon rates, it was shown in the previous chapters that the grey values in the projection can be modelled using a Normal distribution. Thus, $\text{projection}_{x,y}$ is a random quantity. The simulation $\emph{aRTist}_{x,y}$ is not random because this was obtained through a computer software. Given the 19 selected projections used for training the variance-mean model, the variance prediction $s^2\left[\emph{aRTist}_{x,y}\right]$ is not random because the variance-mean model was fitted before encountering the remaining projection to be compared with \emph{aRTist}. The variance-mean model can be made random if a different set of 19 projections was used to train the model each time this inference was conducted, but this shall not be considered here.

If there are no significant differences between the obtained and \emph{aRTist} projections, then the randomness of the test statistics can be approximately quantified as
\begin{equation}
Z_{x,y}\sim \normal(0,1) \ .
\end{equation}
As with usual statistical convention, an upper case $Z$ denotes a random variable. A lower case $z$ denotes a realisation or an observation of that random variable.

@@

It can be noted that this mode finding method can be aplied to other distributions in general.

@@

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$, for example, a quadratic term can be added to the GLM fit. The method used in fitting the smoothing spline could contribute to the underfitting. For example, if the smoothing spline was not smooth enough so that multiple local minima appeared, it would add variance to the optimal bandwidth if the optimisation algorithm converges to a local minimum. A basic linear relationship was chosen to avoid overfitting and to keep the rule of thumb simple. It would also choose a large bandwidth for small $n$ which would improve the performance $\widehat{\mu}_0$ by working on a smoother density estimate for the price of a less optimal $s^2\left[\emph{aRTist}_{x,y}\right]$.

@@

The empirical null mean is effectively the mode and the empirical null standard deivation is a measure of dispersion. It was investigated how they compare with various estimates of central tendency and dispersion for varying sample sizes $n$. The empirical sampling distribution of these estimators was observed to look for properties such as bias, variance and robustness.

The central tendency estimates looked at were the median, the mean and Tukey's biweighted mean \citep{beaton1974fitting}. In summary, the biweighted mean weigh each data point using a smooth function of its residual and ignores any data points with residuals greater than a certain value. \emph{MATLAB}'s \texttt{robustfit} function was used for the biweighted mean and selects any tuning parmaeters.

The dispersion estimates looked at were the interquartile range (IQR) $\div 1.349$, the standard deviation and the median absolute deviation around the mode (MADA-mode) $\times 1.483$, . The normalisation constants were used so that MADA-mode and IQR can be compared directly to the standard deviation. MADA-mode was considered to investigate how estimators, which depends on the mode, perform.

In the next section, these estimators are used on a circular kernel with radius $r$ containing $\pi r^2$ pixels. To make results comparable, the sample size $n$ may be represented using $r=\sqrt{n/\pi}$.

A sample of standard Normal random variables was simulated to obtain estimates of central tendency and dispersion. This was repeated 100 times to obtain a sampling distribution. The results are shown in Figures \ref{fig:inference_NullIidMean} and \ref{fig:inference_NullIidStd}. The mode and empirical null standard deviation both suffered from high variance, much higher than typical robust estimators.

@@

It is interesting to compare the empirical null standard devaition with MADA-mode, in Figure \ref{fig:inference_NullIidStd}, because they both depend on the mode. The mode suffered from high variance with may affect dispersion estimators using the mode. The figure shows that MADA-mode has less variance than the empirical null standard deviation. This suggests that using the curvature of the density estimate at the mode is a main source of variance for the empirical null standard devation.

To test for robustness, a sample of mixture of Gaussian random variables were simulated with the following distribution: $\normal(0,1)$ with probability $90\%$, $\normal(3,1)$ with probability $10\%$. The $\normal(3,1)$ component acts as non-null statistics and robust estimators should be unaffected by it. The results are shown in Figures \ref{fig:inference_NullIidMixtureMean} and \ref{fig:inference_NullIidMixtureStd}. The empirical null estimators are robust to the non-null component but still suffers from high variance. The rest of the estimators were affected by the non-null statistics, making them unsuitable to estimate parameters of the null parameters.

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean2.eps}
        \caption{Bisquare weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean3.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean4.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd3.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd4.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidStd}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean2.eps}
        \caption{Bisquared weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean3.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean4.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMixtureMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd3.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd4.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidMixtureStd}
\end{figure}

@@

The empirical null filter extends the empirical null to images. The parameters of the null distribution varies spatially, slowly and smoothly. Let the test statistic of the pixel at position $(x,y)$ be
\begin{equation}
Z_{x,y}\sim\normal(\mu_{0,x,y},\widehat{\sigma}_{0,x,y}^2)
\end{equation}
for $x=1,2,\dotsc,W$ and $y=1,2,\dotsc,H$. $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ are the null mean and null standard deviation at position $(x,y)$ respectively. Define the null hypotheses to be
\begin{equation}
  H_{0,x,y}:\mu_{0,x,y}=\widehat{\mu}_{0,x,y}
\end{equation}
which are tested against
\begin{equation}
  H_{1,x,y}:\mu_{0,x,y}\neq\widehat{\mu}_{0,x,y} \ .
\end{equation}
 The empirical null filter aims to estimate the parameters $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ for all $x$ and $y$.

@@

The parameter of choice is the kernel radius $r$. A radius too small will treat defects as the null and will not detect them, too big can also cause problems as well. The empirical null filter assumes that the null parameters vary smoothly and slowly so that the null test statistics captured by the kernel is somewhat Normal. If the radius is too big, the kernel will capture distanced test statistics with very different null distributions. If all of these test statistics are pulled together into a histogram, the null distribution may not look Normal at all which can cause problems when fitting a Normal null distribution. Furthermore, if $r\rightarrow \infty$, then this is just using the empirical null on all pixels.

The radius controls the proportion of pixels which are null captured by the kernel, denoted as $\pi_0$. It should be selected so that $\pi_0$ is of sensible value, for example, $>90\%$. In practice, the defects are unknown so various kernel radiuses should be tried out. The kernel radius should be much larger than the size of the defect.

@@

An experiment was conducted to demonstrate the effects of the empirical null filter on a simulated $\SI{256}{\pixel}\times\SI{256}{\pixel}$ Gaussian image, where all the pixels have value distributed as standard Normal. The purpose is to simulate test statistics which are null and few positive results, subject to the FDR, should be detected. The test statistics after filtering should preserve its original distribution, or at least close to it. The statistical moments before and after filtering should be the same and were investigated. Because the empirical null mean and empirical null standard deviation are random variables, the filtered pixels will never be Normal and any normality tests would be too strict for this experiment.

Contamination is defined as some smooth and slowly varying function added and/or multiplied to an image. Conducting hypotheses testing on a contaminated Gaussian image may cause errors if the null distributions were incorrectly specified as standard Normal. In this experiment, the contamination was a gradient such that the pixel at position $(x,y)$ has a value distributed as $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_{0,x,y}^2)$ where
\begin{equation}
  \mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0) \ ,
\end{equation}
$(x_0,y_0)$ is the centre of the image and $\sigma_{0,x,y}=2$ for all $(x,y)$.

A caveat is that some bias to the null standard deviation estimation would be introduced because of the contamination. This is because sources of variance captured by the kernel are from $\sigma_{0,x,y}$ and also from the variability of $\mu_{0,x,y}$. This can be shown with an example. Let $Z_{x,y}$ be the value of the pixels captured by the circular kernel, centred at the origin, for integer values of $(x,y)$ such that $x^2+y^2\leqslant r^2$. Suppose that $Z_{x,y}\sim\normal(ax+by,\sigma^2)$. The quantities of interest are the expectation and variance of all the values contained in the kernel because this is the quantity the estimators are estimating.

The calculations of the expectation and variance of $Z_{x,y}$ can be approximated by treating $x$ and $y$ as uniformly distributed within a circle centred at the origin with radius $r$. The expectation is
\begin{equation}
\expectation[Z_{X,Y}] = \expectation\expectation[Z_{X,Y}|X,Y]=0 \ .
\end{equation}

The variance is given as $\variance[Z_{X,Y}] = \expectation\variance[Z_{X,Y}|X,Y] + \variance\expectation[Z_{X,Y}|X,Y]$. From the distribution of $Z_{X,Y}$, $\expectation\variance[Z_{X,Y}|X,Y] = \sigma^2$ and $\variance\expectation[Z_{X,Y}|X,Y]=\variance[aX+bY]$. This is worked out to be
\begin{align}
\variance[aX+bY] &= \dfrac{1}{\pi r^2}\int_{\rho=0}^{\rho=r}\int_{\theta=0}^{\theta=2\pi}
\rho^3(a\cos\theta + b\sin\theta)^2 \diff \theta \diff \rho
\nonumber\\
& = \dfrac{1}{4}(a^2+b^2)r^2
\end{align}
thus
\begin{equation}
\variance[Z_{X,Y}] = \dfrac{1}{4}(a^2+b^2)r^2 + \sigma^2 \ .
\end{equation}
This shows that when estimating the null variance on a contaminated Gaussian image, there may be a $(a^2+b^2)r^2/4$ bias. If the contamination is slowly varying, that is $a$ and $b$ are small, then the bias should be small.

@@

For a given kernel radius $r$, 100 different Gaussian images were filtered to investigate the within image mean, standard deviation and kurtosis of the filtered statistics. Various other filters for normalisation were used as well such as the MADA-mode null filter, median IQR null filter and the mean variance null filter. They use different estimators to estimate the null parameters as their name suggests, Table \ref{table:inference_nullFilters} clarifies them.

\begin{table}
  \centering
  \begin{tabular}{l|l|l}
    Filters&Null mean&Null standard deviation\\\hline
    MADA-mode null&Mode&MADA-mode $\times1.483$\\
    Median IQR null&Median&IQR $\div1.349$\\
    Mean variance null&Mean&Standard deviation
  \end{tabular}
  \caption{Various filters using different estimators to estimate the null parameters are described here.}
  \label{table:inference_nullFilters}
\end{table}

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_mean.eps}
    \caption{Empirical null filter}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_mean.eps}
    \caption{MADA-mode null filter}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_mean.eps}
    \caption{Median IQR null filter}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_mean.eps}
    \caption{Mean var null filter}
  \end{subfigure}
  }
  \caption{The within image sample mean of a filtered Gaussian image of size $\SI{256}{\pixel}\times\SI{256}{\pixel}$.  The boxplots summarise the 100 repeated simulations of the image. The dashed lines show the 95\% confidence interval using standard tests and assuming independence.}
  \label{fig:inference_AllNullGaussian_mean}
\end{figure}

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_variance.eps}
    \caption{Empirical null filter}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_variance.eps}
    \caption{MADA-mode null filter}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_variance.eps}
    \caption{Median IQR null filter}
  \end{subfigure}
  \begin{subfigure}[b]{\subSize}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_variance.eps}
    \caption{Mean var null filter}
  \end{subfigure}
  }
  \caption{The within image sample standard deviation of a filtered Gaussian image of size $\SI{256}{\pixel}\times\SI{256}{\pixel}$. The boxplots summarise the 100 repeated simulations of the image. The dashed lines show the 95\% confidence interval using standard tests and assuming independence.}
  \label{fig:inference_AllNullGaussian_var}
\end{figure}
